---
title: "AM14_GA1"
author: "Group 10"
date: "1/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 

```{r, Load libraries}
library(ggplot2)  
library(dplyr)
library(tidyverse)
library(readxl)
library(lubridate)
library(skimr)
library(moments)
```

## Question 1

Load the monthly stock data. We will also load the daily stock data, which we'll need later in the assignment.

```{r, Load the data, warning=FALSE}
# Read in the monthly data using the 'read_excel' function: 
monthly_data <- read_excel("PS1_Monthly.xlsx") %>%
  # Perform an initial cleaning of the names in the dataset.
  janitor::clean_names() %>%  
  # Mutate and transform dates stored in character/numeric vectors to Date objects
  mutate(date = ymd(date))

# Skim the data - this allows us to view specific summary statistics
skimr::skim(monthly_data)
```


The "PS1_Monthly" data contains information related to the stocks for the following 
unique companies. 

```{r Question 1 continued}
# Output all the unique company names
print(unique(monthly_data$comnam))
```






## Question 2

In this section, we use the holding period returns to create a total return index 
for MSFT + GE stocks, as well as the S&P500 index. We compare the returns 
where dividends are reinvested, against the returns without dividends for 
each of these stocks/indices.


```{r, Question 2}
# Create variable 'question2' which contains the relevant stocks
question2 <- c('MSFT', 'GE')

# Create a new variable 'monthly_q2' for determining the returns. 
monthly_q2 <- monthly_data %>%
  # Use 'question2' variable to filter out for 'MSFT' and 'GE'
  filter(ticker %in% question2) %>%
  # Group the two 'tickers' above
  group_by(ticker) %>% 
  # Mutate new columns for the returns (+1)
  mutate(ret1 = ret+1,
         retx1 = retx +1,
         sprtrn1 = sprtrn +1) %>% 
  # Mutate for the cumulative product of the previous variables to work out the 
  # the inverse returns - inv_ret accounts for  stock returns with dividends, 
  # inv_retx accounts for stock returns without dividends, whilst inv_sprtrn
  # accounts for S&P500 returns. 
  mutate(inv_ret = cumprod(ret1),
         inv_retx = cumprod(retx1),
         inv_sprtrn = cumprod(sprtrn1)) %>% 
  # Reverse the inverse for these variables
  select(-ret1,-retx1,-sprtrn1)


# Call monthly_q2 variable for plotting - we first plot MSFT
monthly_q2 %>% 
  # Filter for MSFT
  filter(ticker == "MSFT") %>% 
  # Initialise ggplot
  ggplot() +
  # Plot lines for MSFT returns with dividends and without dividends
  geom_line(aes(x = date, y=inv_ret, color="With dividend"), alpha = 0.5, size = 1)+
  geom_line(aes(x = date, y=inv_retx, color="Without dividend"),alpha = 0.5, size = 1)+
  # Add a line plot for the S&P500 returns
  geom_line(aes(x = date, y=inv_sprtrn, color="S&P 500"),alpha = 0.5, size = 1)+
  # Set theme for the plot
  theme_minimal()+
  # Assign labels
  labs(title="Investment with/without dividends for MSFT",
    y="Investment return",
    x="Date")+
  # Assign colours to the keys
  scale_color_manual(name = "Y series", values = c("Without dividend" = "blue", "With dividend" = "red","S&P 500" = "green"))


# Call monthly_q2 variable for plotting GE
monthly_q2 %>% filter(ticker == "GE") %>% 
  # Initialise ggplot
  ggplot() +
  # Use the geom_line function to plot GE returns with and without dividends
  geom_line(aes(x = date, y=inv_ret, color="With dividend"), alpha = 0.5, size = 1)+
  geom_line(aes(x = date, y=inv_retx, color="Without dividend"),alpha = 0.5, size = 1)+
  # Create a line plot of the S&P500 for comparison
  geom_line(aes(x = date, y=inv_sprtrn, color="S&P 500"),alpha = 0.5, size = 1)+
  # Set theme for the plot
  theme_minimal()+
  # Apply lables to this second plot
  labs(title="Investment with/without dividends for GE",
    y="Investment return",
    x="Date")+
  # Assign colours to corresponding lines
  scale_color_manual(name = "Y series", values = c("Without dividend" = "blue", "With dividend" = "red","S&P 500" = "green"))

```

As we can see, the peaks and troughs for all plots are somewhat aligned. 
This shows that all of these stocks/indices follow similar behavioral 
+ investing trends. As expected, we can also see that the investment 
returns **with** dividends will be higher than those **without** dividends.



## Question 3

Here, we generate a new variable that contains log returns (LRET) - we calculate 
sample statistics for these log returns and then plot the normal returns against 
the log returns for "MSFT". 

```{r Question 3}
monthly_log <- monthly_data %>% 
  # Create a log return column 'lret' which takes the log of the return
  mutate(lret = log(1+ret)) 

# Create a new dataframe for the summary statistics we calculate
sum_table_monthly <- monthly_log %>% 
  # Group by the 'ticker' so we can calculate statistics for each one
  group_by(ticker) %>% 
  # Use the summarise function to calculate different sample statistics
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))

# Output the new calculated sample statistics 
sum_table_monthly

monthly_log %>% 
  # Filter out "MSFT" from our 'monthly log' dataframe so that we can plot
  filter(ticker == "MSFT") %>%
  # Call ggplot
  ggplot() +
  # Create line plots for both log returns and normal returns - differentiated by colour
  geom_line(aes(x= date, y=ret, color="ret"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=lret, color="lret"), alpha = 0.5, size = 1)+
  # Set the figure theme
  theme_minimal()+
  # Apply labels to the plot
  labs(title="Ordinary Return v.s. Log Return for MSFT",
    y="Return",
    x="Date")+
  # Assign the appropraite colours to the plot and its legend
  scale_color_manual(name = "Y series", values = c("ret" = "blue", "lret" = "red"))
  
```

As, we can see from the summary statistics, taking the log return seems to 
exaggerate the key statistics (mean, variance, skew, etc.) associated with the 
returns. This can be highlighted clearly on the plot which compares the normal
returns with the log returns for "MSFT". 

Clearly, we can see that there is a significant difference between the ordinary
returns and the log returns over time. This points towards the returns being 
non-negligible, with relatively high percentage changes. Ordinary returns and log
returns are only equal when they are **zero**, and they are approximately equal 
when they remain small. Therefore, the difference between the ordinary and the log
returns highlights that these returns are non-negligible - and that there is 
significant percentage change (which points towards relatively high volatility) 
for "MSFT".




# Question 4

In this section we download the data from the PS1_Daily.xlsx file - this contains
two separate sheets - the first 'HPR_daily" is **downloaded below** and contains 
the daily holding period for the 6 stocks.

```{r}
daily_data <- read_excel("PS1_Daily.xlsx", sheet = "HPR_daily", skip = 1) %>%
  janitor::clean_names()%>% 
  mutate(date = ymd(date))
skimr::skim(daily_data)
```

```{r, Question 5}
question5 <- c('msft', 'ge',"sprtrn")

daily_q5 <- daily_data %>% pivot_longer(!date,names_to = "ticker",values_to = "ret") %>%
  filter(ticker %in% question5) %>%
  group_by(ticker) %>% 
  arrange(ticker,date) %>% 
  mutate(ret1 = ret+1) %>% 
  mutate(inv_ret = cumprod(ret1)) %>% 
  select(-ret1)

daily_q5 %>% 
  ggplot(aes(x = date, y=inv_ret, group=ticker, color=ticker)) +
  geom_line()+
  theme_minimal()+
  labs(
    title="Daily Investment",
    y="Investment return",
    x="Date"
  )

```


```{r q6}
daily_log <-daily_data %>% pivot_longer(!date,names_to = "ticker",values_to = "ret") %>% 
  mutate(lret = log(1+ret))

sum_table_daily <-  daily_log %>% 
  group_by(ticker) %>% 
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))
sum_table_daily
```




## Question 7

```{r q7-1}
#MSFT
#Monthly log returns
monthly_log %>%
  filter(ticker == "MSFT") %>%
  select(lret) %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(monthly_log$lret),
                            sd = sd(monthly_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "Microsoft - Histogram of Log Monthly Returns",
    x = "Log Monthly Returns",
    y = "Density"
  )

#Daily log returns
daily_log %>%
  filter(ticker == "msft") %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(daily_log$lret),
                            sd = sd(daily_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "Microsoft -  Histogram of Log Daily Returns",
    x = "Log Daily Returns",
    y = "Density"
  )

# While both graphs are leptokurtic (more values centered around the mean; higher bell)
# daily log returns by far have the highest kurtosis
```


```{r q7-2}
#General Electric
monthly_log %>%
  filter(ticker == "GE") %>%
  select(lret) %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(monthly_log$lret),
                            sd = sd(monthly_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "General Electric - Histogram of Log Monthly Returns",
    x = "Log Monthly Returns",
    y = "Density"
  )

#Daily log returns
daily_log %>%
  filter(ticker == "ge") %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(daily_log$lret),
                            sd = sd(daily_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "General Electric - Histogram of Log Daily Returns",
    x = "Log Daily Returns",
    y = "Density"
  )
```

# Question 8

Hovik: use MSFT, GE and JPM

```{r}
daily_q8_ret <- daily_data %>%
  select(-c(xom,intc,c,vwretd))  %>%
  pivot_longer(!date,names_to = "ticker",values_to = "ret")%>% 
  mutate(lret = log(1+ret))

daily_q8_inv <-  daily_data %>%
         select(-c(xom,intc,c,vwretd))  %>%
         mutate(msft1 = msft +1,
         inv_msft = msft1 * lag(msft1),
         ge1 = ge +1,
         inv_ge = ge1 * lag(ge1),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = sprtrn1 * lag(sprtrn1),
         jpm1 = jpm +1,
         inv_jpm = jpm1 * lag(jpm1),) %>% 
  mutate(inv_msft = if_else(is.na(inv_msft),1,inv_msft),
         inv_ge = if_else(is.na(inv_ge),1,inv_ge),
         inv_jpm = if_else(is.na(inv_jpm),1,inv_jpm),
         inv_sprtrn = if_else(is.na(inv_sprtrn),1,inv_sprtrn)) %>% 
  select(-msft1,-ge1,-sprtrn1,-jpm1,-msft,-ge,-sprtrn,-jpm) %>%
  mutate(msft = inv_msft, ge = inv_ge, sprtrn = inv_sprtrn, jpm = inv_jpm) %>%
  select(-inv_msft,-inv_ge,-inv_sprtrn,-inv_jpm) %>%
  pivot_longer(!date,names_to = "ticker",values_to = "inv")
  
daily_q8 <- left_join(daily_q8_ret, daily_q8_inv)
daily_q8
```

# Question 9
```{r, covariance matrix for log return series}

#Covariance matrix for log returns
covariance_data_log <- daily_data %>%
  mutate(
    msft_log_ret = log(1 + msft),
    xom_log_ret = log(1 + xom),
    ge_log_ret = log(1 + ge),
    jpm_log_ret = log(1 + jpm),
    intc_log_ret = log(1 + intc),
    c_log_ret = log(1 + c),
    )

cov(covariance_data_log %>% 
      select(msft_log_ret, xom_log_ret, ge_log_ret, jpm_log_ret, intc_log_ret, c_log_ret))


#Covariance matrix for returns

cov(covariance_data_log %>% 
      select(msft, xom, ge, jpm, intc, c))


#Covariance matrix for squared returns
covariance_data_sq_ret <- daily_data %>%
  mutate(
    msft_sq_ret = msft**2,
    xom_sq_ret = xom**2,
    ge_sq_ret = ge**2,
    jpm_sq_ret = jpm**2,
    intc_sq_ret = intc**2,
    c_sq_ret = c**2,
    )

cov(covariance_data_sq_ret %>% 
      select(msft_sq_ret, xom_sq_ret, ge_sq_ret, jpm_sq_ret, intc_sq_ret, c_sq_ret))
```

# Q10

```{r, ACF function}
#Creating a series for returns
acf_data <- daily_data %>% 
  pivot_longer(!date,names_to = "ticker",values_to = "ret")

#ACF function for returns
acf_data %>% 
  select(ret) %>% 
  acf()

#ACF for returns squared
acf_data %>% 
  mutate(returns_squared = ret ** 2) %>% 
  select(returns_squared) %>% 
  acf()

#ACF for absolute returns
acf_data %>% 
  mutate(abs_returns = abs(ret)) %>% 
  select(abs_returns) %>% 
  acf()
```



# Q11

```{r, portfolio}
#Selecting the 3 stocks
portfolio_data <- daily_data %>%
  select(date, msft, ge, jpm)

#Assigning weights arbitrarily
w_msft = 0.4
w_ge = 0.2
w_jpm = 0.4

#Portfolio
portfolio_data <- portfolio_data %>% 
  mutate(
    portfolio_returns = 1 + ((w_msft * msft) + (w_ge * ge) + (w_jpm * jpm))
  )
```


# Q12

```{r, investment growth}
#Investment growth of $1
portfolio_data <- portfolio_data %>% 
  mutate(
    #Cumulative product of returns
    investment_growth = cumprod(portfolio_returns)
  )


#Plot of investment growth
portfolio_data %>% 
  ggplot() +
  geom_line(aes(x= date, y=investment_growth)) +
  theme_minimal()+
  labs(
    title="Investment Growth for Portfolio",
    y="Investment return",
    x="Date"
  )
```



```{r comparison of portfolio return with individual stocks}
return_data <- daily_data %>%
  #Selecting relevant columns
  select(c(date, msft, jpm, ge))  %>%
  
  #Calculating returns
  mutate(msft_ret = msft + 1,
         inv_msft = cumprod(msft_ret),
         
         ge_ret = ge +1,
         inv_ge = cumprod(ge_ret),
         
         jpm_ret = jpm +1,
         inv_jpm = cumprod(jpm_ret)) %>% 
  select(date, inv_msft, inv_ge, inv_jpm)

#Joining data of individual stocks with portfolio
comparison_data <- return_data %>% 
  inner_join(portfolio_data %>% select(date, investment_growth), by = "date")

#Plotting returns for all
comparison_data %>% 
  ggplot() +
  geom_line(aes(x= date, y=inv_msft, color="msft"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_jpm, color="jpm"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_ge, color="ge"),alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=investment_growth, color="portfolio"),alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Comparison for investment growth of portfolio vs individual stocks",
    y="Investment return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("msft" = "blue", "ge" = "green","jpm"="red", "portfolio"="orange"))
```


