---
title: "AM14_GA1"
author: "Group 10"
date: "1/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 

```{r, Load libraries}
library(ggplot2)  
library(dplyr)
library(tidyverse)
library(readxl)
library(lubridate)
library(skimr)
library(moments)
```

## Question 1

Load the monthly stock data. We will also load the daily stock data, which we'll need later in the assignment.

```{r, Load the data, warning=FALSE}
monthly_data <- read_excel("PS1_Monthly.xlsx") %>%
  janitor::clean_names() %>%  
  mutate(date = ymd(date))

skimr::skim(monthly_data)

```


## Question 2
Yuting: Not sure

Now, returns bla bla bla 

Also - upon inspection, the dates are not in the right format and need to be transformed. 

```{r, Question 2}
question2 <- c('MSFT', 'GE')

monthly_q2 <- monthly_data %>%
  filter(ticker %in% question2) %>% 
  mutate(ret1 = ret +1,
         inv_ret = if_else(ret1 == 1, 1, ret1 * lag(ret1)),
         retx1 = retx +1,
         inv_retx = if_else(retx1 == 1, 1, retx1 * lag(retx1)),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = if_else(sprtrn1 == 1, 1, sprtrn1 * lag(sprtrn1))) %>% 
  select(-ret1,-retx1,-sprtrn1)

monthly_q2 %>% filter(ticker == "MSFT") %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y=inv_ret), color="red", alpha = 0.5, size = 1)+
  geom_line(aes(y=inv_retx), color="blue",alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment without dividends for MSFT",
    y="Investment return",
    x="Date"
  )

monthly_q2 %>% filter(ticker == "GE") %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y=inv_ret), color="red", alpha = 0.5, size = 1)+
  geom_line(aes(y=inv_retx), color="blue",alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment without dividends for GE",
    y="Investment return",
    x="Date"
  )
```

## Question 3
yuting: since some return < 0, I filter out them

```{r}
monthly_log <- monthly_data %>% filter(ret > 0) %>% mutate(lret = log(ret)) 
sum_table_monthly <- monthly_log %>% 
  group_by(ticker) %>% 
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))
sum_table_monthly

monthly_log %>% filter(ticker == "MSFT") %>%
  ggplot() +
  geom_line(aes(x= date, y=ret, color="ret"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=lret, color="lret"), alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Return v.s. log Return for MSFT",
    y="Return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("ret" = "blue", "lret" = "red"))
  
  
```
# Q4
Yuting: I manually separate two worksheet into two excel files.

```{r}
daily_data <- read_excel("PS1_Daily_HPR_daily.xlsx") %>%
  janitor::clean_names()%>% 
  mutate(date = ymd(date))
skimr::skim(daily_data)
```

```{r, Question 5, fig.width=12}
daily_q5 <- daily_data %>%
  select(-c(xom,jpm,intc,c,vwretd))  %>%
  mutate(msft1 = msft +1,
         inv_msft = msft1 * lag(msft1),
         ge1 = ge +1,
         inv_ge = ge1 * lag(ge1),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = sprtrn1 * lag(sprtrn1)) %>% 
  mutate(inv_msft = if_else(is.na(inv_msft),1,inv_msft),
         inv_ge = if_else(is.na(inv_ge),1,inv_ge),
         inv_sprtrn = if_else(is.na(inv_sprtrn),1,inv_sprtrn)) %>% 
  select(-msft1,-ge1,-sprtrn1)

daily_q5 %>% 
  ggplot() +
  geom_line(aes(x= date, y=inv_msft, color="msft"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_sprtrn, color="sprtrn"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_ge, color="ge"),alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment for MSFT, GE, S&P 500",
    y="Investment return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("msft" = "blue", "sprtrn" = "green","ge"="red"))

```

```{r q6}
daily_log <-daily_data %>% pivot_longer(!date,names_to = "ticker",values_to = "ret")%>% 
  filter(ret > 0) %>% mutate(lret = log(ret))

sum_table_daily <-  daily_log %>% 
  group_by(ticker) %>% 
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))
sum_table_daily
```

