---
title: "AM14_GA1"
author: "Group 10"
date: "1/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries 

```{r, Load libraries}
library(ggplot2)  
library(dplyr)
library(tidyverse)
library(readxl)
library(lubridate)
library(skimr)
library(moments)
```

## Question 1

Load the monthly stock data. We will also load the daily stock data, which we'll need later in the assignment.

```{r, Load the data, warning=FALSE}
# Read in the monthly data using the 'read_excel' function: 
monthly_data <- read_excel("PS1_Monthly.xlsx") %>%
  # Perform an initial cleaning of the names in the dataset.
  janitor::clean_names() %>%  
  # Mutate and transform dates stored in character/numeric vectors to Date objects
  mutate(date = ymd(date))

# Skim the data - this allows us to view specific summary statistics
skimr::skim(monthly_data)
```


The "PS1_Monthly" data contains information related to the stocks for the following 
unique companies. 

```{r}
print(unique(monthly_data$comnam))
```





## Question 2

Yuting: Not sure

Now, returns bla bla bla 

Also - upon inspection, the dates are not in the right format and need to be transformed. 

```{r, Question 2}
question2 <- c('MSFT', 'GE')

monthly_q2 <- monthly_data %>%
  filter(ticker %in% question2) %>% 
  mutate(ret1 = ret +1,
         inv_ret = if_else(ret1 == 1, 1, ret1 * lag(ret1)),
         retx1 = retx +1,
         inv_retx = if_else(retx1 == 1, 1, retx1 * lag(retx1)),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = if_else(sprtrn1 == 1, 1, sprtrn1 * lag(sprtrn1))) %>% 
  select(-ret1,-retx1,-sprtrn1)

monthly_q2 %>% filter(ticker == "MSFT") %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y=inv_ret), color="red", alpha = 0.5, size = 1)+
  geom_line(aes(y=inv_retx), color="blue",alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment without dividends for MSFT",
    y="Investment return",
    x="Date"
  )

monthly_q2 %>% filter(ticker == "GE") %>% 
  ggplot(aes(x=date)) +
  geom_line(aes(y=inv_ret), color="red", alpha = 0.5, size = 1)+
  geom_line(aes(y=inv_retx), color="blue",alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment without dividends for GE",
    y="Investment return",
    x="Date"
  )
```




## Question 3
yuting: since some return < 0, I filter out them

```{r}
monthly_log <- monthly_data %>% filter(ret > 0) %>% mutate(lret = log(ret)) 
sum_table_monthly <- monthly_log %>% 
  group_by(ticker) %>% 
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))
sum_table_monthly

monthly_log %>% filter(ticker == "MSFT") %>%
  ggplot() +
  geom_line(aes(x= date, y=ret, color="ret"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=lret, color="lret"), alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Return v.s. log Return for MSFT",
    y="Return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("ret" = "blue", "lret" = "red"))
  
  
```

# Q4
Yuting: I manually separate two worksheet into two excel files.

```{r}
daily_data <- read_excel("PS1_Daily.xlsx", sheet = "HPR_daily", skip = 1) %>%
  janitor::clean_names()%>% 
  mutate(date = ymd(date))
skimr::skim(daily_data)
```

```{r, Question 5, fig.width=12}
daily_q5 <- daily_data %>%
  select(-c(xom,jpm,intc,c,vwretd))  %>%
  mutate(msft1 = msft +1,
         inv_msft = msft1 * lag(msft1),
         ge1 = ge +1,
         inv_ge = ge1 * lag(ge1),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = sprtrn1 * lag(sprtrn1)) %>% 
  mutate(inv_msft = if_else(is.na(inv_msft),1,inv_msft),
         inv_ge = if_else(is.na(inv_ge),1,inv_ge),
         inv_sprtrn = if_else(is.na(inv_sprtrn),1,inv_sprtrn)) %>% 
  select(-msft1,-ge1,-sprtrn1)

daily_q5 %>% 
  ggplot() +
  geom_line(aes(x= date, y=inv_msft, color="msft"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_sprtrn, color="sprtrn"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_ge, color="ge"),alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Investment for MSFT, GE, S&P 500",
    y="Investment return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("msft" = "blue", "sprtrn" = "green","ge"="red"))

```


```{r q6}
daily_log <-daily_data %>% pivot_longer(!date,names_to = "ticker",values_to = "ret")%>% 
  filter(ret > 0) %>% mutate(lret = log(ret))

sum_table_daily <-  daily_log %>% 
  group_by(ticker) %>% 
  summarise(ret_mean = mean(ret),
            logret_mean = mean(lret, na.rm = TRUE),
            ret_var = var(ret),
            logret_var = var(lret, na.rm = TRUE),
            ret_skew = skewness(ret),
            logret_skew = skewness(lret, na.rm = TRUE),
            ret_kur = kurtosis(ret),
            logret_kur = kurtosis(lret, na.rm = TRUE))
sum_table_daily
```

## Question 7

```{r q7-1}
#MSFT
#Monthly log returns
monthly_log %>%
  filter(ticker == "MSFT") %>%
  select(lret) %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(monthly_log$lret),
                            sd = sd(monthly_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "Microsoft - Histogram of Log Monthly Returns",
    x = "Log Monthly Returns",
    y = "Density"
  )

#Daily log returns
daily_log %>%
  filter(ticker == "msft") %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(daily_log$lret),
                            sd = sd(daily_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "Microsoft -  Histogram of Log Daily Returns",
    x = "Log Daily Returns",
    y = "Density"
  )

# While both graphs are leptokurtic (more values centered around the mean; higher bell)
# daily log returns by far have the highest kurtosis
```


```{r q7-2}
#General Electric
monthly_log %>%
  filter(ticker == "GE") %>%
  select(lret) %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(monthly_log$lret),
                            sd = sd(monthly_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "General Electric - Histogram of Log Monthly Returns",
    x = "Log Monthly Returns",
    y = "Density"
  )

#Daily log returns
daily_log %>%
  filter(ticker == "ge") %>%
  ggplot(aes(x = lret)) +
  geom_histogram(binwidth = 0.01) +
  stat_function(fun = dnorm,
                args = list(mean = mean(daily_log$lret),
                            sd = sd(daily_log$lret)),
                col = "#1b98e0",
                size = 1) +
  theme_bw() +
  labs(
    title = "General Electric - Histogram of Log Daily Returns",
    x = "Log Daily Returns",
    y = "Density"
  )
```

# Question 8

Hovik: use MSFT, GE and JPM

```{r}
daily_q8_ret <- daily_data %>%
  select(-c(xom,intc,c,vwretd))  %>%
  pivot_longer(!date,names_to = "ticker",values_to = "ret")%>% 
  filter(ret > 0) %>% 
  mutate(lret = log(ret))

daily_q8_inv <-  daily_data %>%
         select(-c(xom,intc,c,vwretd))  %>%
         mutate(msft1 = msft +1,
         inv_msft = msft1 * lag(msft1),
         ge1 = ge +1,
         inv_ge = ge1 * lag(ge1),
         sprtrn1 = sprtrn +1,
         inv_sprtrn = sprtrn1 * lag(sprtrn1),
         jpm1 = jpm +1,
         inv_jpm = jpm1 * lag(jpm1),) %>% 
  mutate(inv_msft = if_else(is.na(inv_msft),1,inv_msft),
         inv_ge = if_else(is.na(inv_ge),1,inv_ge),
         inv_jpm = if_else(is.na(inv_jpm),1,inv_jpm),
         inv_sprtrn = if_else(is.na(inv_sprtrn),1,inv_sprtrn)) %>% 
  select(-msft1,-ge1,-sprtrn1,-jpm1,-msft,-ge,-sprtrn,-jpm) %>%
  mutate(msft = inv_msft, ge = inv_ge, sprtrn = inv_sprtrn, jpm = inv_jpm) %>%
  select(-inv_msft,-inv_ge,-inv_sprtrn,-inv_jpm) %>%
  pivot_longer(!date,names_to = "ticker",values_to = "inv")
  
daily_q8 <- left_join(daily_q8_ret, daily_q8_inv)
daily_q8
```

# Question 9

```{r, covariance matrix for log return series}

#Covariance matrix for log returns
covariance_data_log <- daily_data %>%
  mutate(
    msft_log_ret = log(1 + msft),
    xom_log_ret = log(1 + xom),
    ge_log_ret = log(1 + ge),
    jpm_log_ret = log(1 + jpm),
    intc_log_ret = log(1 + intc),
    c_log_ret = log(1 + c),
    )

cov(covariance_data_log %>% 
      select(msft_log_ret, xom_log_ret, ge_log_ret, jpm_log_ret, intc_log_ret, c_log_ret))


#Covariance matrix for returns

cov(covariance_data_log %>% 
      select(msft, xom, ge, jpm, intc, c))


#Covariance matrix for squared returns
covariance_data_sq_ret <- daily_data %>%
  mutate(
    msft_sq_ret = msft**2,
    xom_sq_ret = xom**2,
    ge_sq_ret = ge**2,
    jpm_sq_ret = jpm**2,
    intc_sq_ret = intc**2,
    c_sq_ret = c**2,
    )

cov(covariance_data_sq_ret %>% 
      select(msft_sq_ret, xom_sq_ret, ge_sq_ret, jpm_sq_ret, intc_sq_ret, c_sq_ret))
```

# Q10

```{r, ACF function}
#Creating a series for returns
acf_data <- daily_data %>% 
  pivot_longer(!date,names_to = "ticker",values_to = "ret")

#ACF function for returns
acf_data %>% 
  select(ret) %>% 
  acf()

#ACF for returns squared
acf_data %>% 
  mutate(returns_squared = ret ** 2) %>% 
  select(returns_squared) %>% 
  acf()

#ACF for absolute returns
acf_data %>% 
  mutate(abs_returns = abs(ret)) %>% 
  select(abs_returns) %>% 
  acf()
```



# Q11

```{r, portfolio}
#Selecting the 3 stocks
portfolio_data <- daily_data %>%
  select(date, msft, ge, jpm)

#Assigning weights arbitrarily
w_msft = 0.4
w_ge = 0.2
w_jpm = 0.4

#Portfolio
portfolio_data <- portfolio_data %>% 
  mutate(
    portfolio_returns = 1 + ((w_msft * msft) + (w_ge * ge) + (w_jpm * jpm))
  )
```


# Q12

```{r, investment growth}
#Investment growth of $1
portfolio_data <- portfolio_data %>% 
  mutate(
    #Cumulative product of returns
    investment_growth = cumprod(portfolio_returns)
  )


#Plot of investment growth
portfolio_data %>% 
  ggplot() +
  geom_line(aes(x= date, y=investment_growth)) +
  theme_minimal()+
  labs(
    title="Investment Growth for Portfolio",
    y="Investment return",
    x="Date"
  )
```



```{r comparison of portfolio return with individual stocks}
return_data <- daily_data %>%
  #Selecting relevant columns
  select(c(date, msft, jpm, ge))  %>%
  
  #Calculating returns
  mutate(msft_ret = msft + 1,
         inv_msft = cumprod(msft_ret),
         
         ge_ret = ge +1,
         inv_ge = cumprod(ge_ret),
         
         jpm_ret = jpm +1,
         inv_jpm = cumprod(jpm_ret)) %>% 
  select(date, inv_msft, inv_ge, inv_jpm)

#Joining data of individual stocks with portfolio
comparison_data <- return_data %>% 
  inner_join(portfolio_data %>% select(date, investment_growth), by = "date")

#Plotting returns for all
comparison_data %>% 
  ggplot() +
  geom_line(aes(x= date, y=inv_msft, color="msft"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_jpm, color="jpm"), alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=inv_ge, color="ge"),alpha = 0.5, size = 1)+
  geom_line(aes(x= date, y=investment_growth, color="portfolio"),alpha = 0.5, size = 1)+
  theme_minimal()+
  labs(
    title="Comparison for investment growth of portfolio vs individual stocks",
    y="Investment return",
    x="Date"
  )+
  scale_color_manual(name = "Y series", values = c("msft" = "blue", "ge" = "green","jpm"="red", "portfolio"="orange"))
```












